---
title: "Single-Inventory and Multi-Inventory Negative Binomial and Logistic Regression Models, Landslide-Only and Landslide-Precipitation"
author: "Lisa Luna"
date: "8/26/2022"
output: html_document
---

This notebook explains and fits negative binomial and logistic regression models.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Set up: library needed packages and read the data

```{r}
# Load libraries
library(brms) #brms version 2.17.0 was used to fit models
library(bayesplot)
library(dplyr)

```

### Landslide-only models ###

```{r}
# Read the data

datadir <- '../03_Figures/v3/'

#these .csv files are created by the Python Jupyter Notebook "01_LandslideInventoryProcessing.ipynb"

#landslide counts by month and inventory
slidesallmi <- read.csv(paste0(datadir, 'slidesallmi.csv'))


#convert the inventory names to a factor to be able to use them as a level in the multi-level model
slidesallmi$inv <- as.factor(slidesallmi$inventory)

#add a binary event indicator to the dataframe for logistic regression (threshold of 1 or more landslides in a month)

slidesallmi$event <- ifelse(slidesallmi$counts == 0, 0, 1)

```


We train two variants of Bayesian regression models to learn the seasonal pattern of landslide activity at monthly resolution.  We use negative binomial regression to estimate the number of landslides observed in a given month in an inventory (intensity) and logistic regression to estimate the presence or absence of reported landslides (probability).  For each regression, we first fit models to data from each inventory separately (single-inventory models), and then fit a model that learns simultaneously from all inventories (multi-inventory models).

# Negative Binomial

## Single-Inventory Negative Binomial models

Here, a separate model is fit to each inventory. We have observations of counts of landslides in individuals months (e.g. January 2002). These individual month observations are $y_i$. These individual months belong to groups - months (e.g. January 2002 belongs to the group "January").  The model learns one negative binomial distribution for each month, as well as the distribution of intercepts between months. 



```{r}
#Negative Binomial

#separate model for each inventory, grouped by month


fit_nasa_NB <- brm(bf(counts ~ 1 + (1|month) + offset(ln_area), #offset accounts for inventory footprint area
                     shape ~ 1 + (1|month)),
               data = subset(slidesallmi, inventory == 'nasa'),
               family = negbinomial(link = "log", link_shape = "log"),
               warmup = 500,
               iter = 2000,
               chains = 4, cores = 4)

# SLIDOh
fit_slidoh_NB <- brm(bf(counts ~ 1 + (1|month) + offset(ln_area), 
                     shape ~ 1 + (1|month)),
               data = subset(slidesallmi, inventory == 'slidoh'),
               family = negbinomial(link = "log", link_shape = "log"),
               warmup = 500,
               iter = 2000,
               chains = 4, cores = 4)

#SLIDOd
fit_slidod_NB <- brm(bf(counts ~ 1 + (1|month) + offset(ln_area), 
                     shape ~ 1 + (1|month)),
               data = subset(slidesallmi, inventory == 'slidod'),
               family = negbinomial(link = "log", link_shape = "log"),
               warmup = 500,
               iter = 2000,
               chains = 4, cores = 4)

#Seattle
fit_seattle_NB <- brm(bf(counts ~ 1 + (1|month) + offset(ln_area), 
                     shape ~ 1 + (1|month)),
               data = subset(slidesallmi, inventory == 'seattle'),
               family = negbinomial(link = "log", link_shape = "log"),
               warmup = 500,
               iter = 2000,
               control = list(max_treedepth = 15, adapt_delta = 0.99),
               chains = 4, cores = 4)


#WLC
fit_wlc_NB <- brm(bf(counts ~ 1 + (1|month) + offset(ln_area), 
                     shape ~ 1 + (1|month)),
               data = subset(slidesallmi, inventory == 'wlc'),
               family = negbinomial(link = "log", link_shape = "log"),
               warmup = 1000,
               iter = 2500,
               control = list(adapt_delta = 0.99, max_treedepth = 15),
               chains = 4, cores = 4, init = "random")


#Chains for the single inventory WLC model did not converge.  This is likely due to the difficulty that the model has with mostly months with 0 landslides reported and few with thousands.  Single-inventory WLC will be excluded from further analysis.


```
```{r}
#check for chain convergence (Rhat should be close to 1)


# Rhat values - Negative Binomial Models
plot(rhat(fit_nasa_NB)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_seattle_NB)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_slidod_NB)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_slidoh_NB)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_wlc_NB)); abline(h = 1.01, lty = 2, col = "red")


```


```{r}
summary(fit_nasa_NB) 
summary(fit_slidoh_NB)
summary(fit_slidod_NB)
summary(fit_seattle_NB)
summary(fit_wlc_NB)

```


## Multi-level Negative Binomial Model 

Here, a multi-level model is fit to data from all of the inventories.  The data is grouped by month and by inventory.  The model learns one group-level intercept for each month from all the inventories, one group-level intercept for each inventory, and the variation between the months and the inventories in each group.


```{r}

#multi-level model grouped by inventory and month

# Negative Binomial 

fit_all_NB <- brm(bf(counts ~ 1 + (1|month) + (1|inv) + offset(ln_area), 
                     shape ~ 1 + (1|month) + (1|inv)),
               data = slidesallmi,
               family = negbinomial(link = "log", link_shape = "log"),
               warmup = 500,
               iter = 2000,
               control = list(adapt_delta = 0.99, max_treedepth = 15),
               chains = 4, cores = 4)

```


```{r}
#check for chain convergence

summary(fit_all_NB) #Rhat = 1.00

plot(rhat(fit_all_NB)); abline(h = 1.01, lty = 2, col = "red")

mcmc_plot(fit_all_NB, 
          type = "trace")
          #type = 'rhat',
          #type = 'nuts_divergence', 
          #type = "neff")

```

# Logistic Regression 

Besides modeling the number of landslides, we also use logistic regression to model the presence or absence of reported landslides.  Logistic regression relies on a Bernoulli distribution with parameter p expressing the probability of observing at least one landslide.  We fit the logistic regression to data that indicate whether one or more landslides were recorded in an inventory in a given month (e.g. the presence or absence of landslides in January 2002 in the Seattle inventory is one data point)

## Single-inventory logistic models

Here, a separate model is fit to each inventory.  Months are levels in this model, so there is one parameter estimated for each month.


```{r}
#logistic regression 

#separate model for each inventory, grouped by month

#NASA GLC

fit_nasa_logst <- brm(bf(event ~ 1 + (1|month)),
                  data = subset(slidesallmi, inventory == 'nasa'),
                  family = bernoulli(link = "logit"),
                  warmup = 500,
                  iter = 2000,
                  chains = 4, cores = 4)
# SLIDOh
fit_slidoh_logst <- brm(bf(event ~ 1 + (1|month)),
                   data = subset(slidesallmi, inventory == 'slidoh'),
                   family = bernoulli(link = "logit"),
                   warmup = 500,
                   iter = 2000,
                   chains = 4, cores = 4)

#SLIDOd
fit_slidod_logst <- brm(bf(event ~ 1 + (1|month)),
                    data = subset(slidesallmi, inventory == 'slidod'),
                    family = bernoulli(link = "logit"),
                    warmup = 500,
                    iter = 2000,
                    control = list(adapt_delta = 0.95),
                    chains = 4, cores = 4)

#Seattle
fit_seattle_logst <- brm(bf(event ~ 1 + (1|month)),
                      data = subset(slidesallmi, inventory == 'seattle'),
                      family = bernoulli(link = "logit"),
                      warmup = 500,
                      iter = 2000,
                      chains = 4, cores = 4)

#WLC
fit_wlc_logst <- brm(bf(event ~ 1 + (1|month)),
                 data = subset(slidesallmi, inventory == 'wlc'),
                 family = bernoulli(link = "logit"),
                 warmup = 500,
                 iter = 2000,
                 control = list(adapt_delta = 0.99),
                 chains = 4, cores = 4)




```


```{r}
#check chain convergence

# Rhat values - Logistic Models
plot(rhat(fit_nasa_logst)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_seattle_logst)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_slidod_logst)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_slidoh_logst)); abline(h = 1.01, lty = 2, col = "red")
plot(rhat(fit_wlc_logst)); abline(h = 1.01, lty = 2, col = "red")

```


## Multi-inventory Logistic Regression

Here, a multi-level model is fit to data from all of the inventories.  Both inventory and month are levels in the model, meaning that intercepts are estimated for each month and for each inventory. 

```{r}
#multi-inventory logistic regression

fit_all_logst <- brm(bf(event ~ 1 + (1 | month) + (1 | inv)),
                 data = slidesallmi,
                 family = bernoulli(link = "logit"),
                 warmup = 500,
                 iter = 2000,
                 control = list(adapt_delta = 0.95),
                 chains = 4, cores = 4)

```


```{r}
#check chain convergence

plot(rhat(fit_all_logst)); abline(h = 1.01, lty = 2, col = "red")


mcmc_plot(fit_all_logst, 
          type = "trace")
          #type = 'rhat',
          #type = 'nuts_divergence', 
          #type = "neff")

```

### Landslide-precipitation models

Now, we condition both multi-inventory models on the average precipitation over the inventory areas in each month between 1991-2020.  

```{r}

#get data

slidesallmi_mrain <- read.csv(paste0(datadir, 'slidesallmi_mrain.csv'))

#convert the inventory names to a factor to be able to use them as a level in the multi-level model
slidesallmi_mrain$inv <- as.factor(slidesallmi_mrain$inventory)

#add a binary event indicator to the dataframe for logistic regression (threshold of 1 or more landslides in a month)

slidesallmi_mrain$event <- ifelse(slidesallmi_mrain$counts == 0, 0, 1)

#rename mean to avoid confusion 
slidesallmi_mrain <- slidesallmi_mrain %>% rename(mean_rain_sp = mean)

#standardize rainfall data to global mean 
slidesallmi_mrain.scale <- slidesallmi_mrain %>%
                            select(c(month, year, inv, counts, event, ln_area, mean_rain_sp)) %>%
                            #mutate(across(c(mean, ante3), ~scale(.)))
                            mutate_at(c("mean_rain_sp"), ~scale(.) %>% as.vector)
                            

#remove the points without data (points before 1990)
slidesallmi_mrain.scale <- slidesallmi_mrain.scale %>%
                            filter(!is.na(mean_rain_sp))

#make a note of these values to be able to to unscale later

slidesallmi_mrain.mean <- slidesallmi_mrain %>% pull(mean_rain_sp) %>% mean(na.rm = TRUE)
slidesallmi_mrain.sd <- slidesallmi_mrain %>% pull(mean_rain_sp) %>% sd(na.rm = TRUE)


```


```{r}
#logistic regression 

fit.logst.rain.all <- brm(formula = event ~ 1 + mean_rain_sp + (1 + mean_rain_sp | month) + (1 + mean_rain_sp | inv),
                           data = slidesallmi_mrain.scale, 
                           family = bernoulli(link = "logit"), 
                           #prior = logst_priors,
                           warmup = 500, 
                           iter = 2000, 
                           chains = 4, 
                           cores = 4, 
                           control = list(adapt_delta = 0.9))

#negative binomial regression 

fit.NB.rain.all <- brm(bf(counts ~ 1 + mean_rain_sp + offset(ln_area) + (1 + mean_rain_sp | month) + (1 + mean_rain_sp | inv), #should be counts per area, check syntax
                                    shape ~ 1 + (1|month) + (1|inv)), #single variance for now, don't let variance vary with rain, but do let it vary with month
                           data = slidesallmi_mrain.scale, 
                           family = negbinomial(link = "log", link_shape = "log"), 
                           #prior = logst_priors,
                           warmup = 1000, 
                           iter = 2500, 
                           chains = 4, 
                           cores = 4,
                          control = list(adapt_delta = 0.99))


```


```{r}
mcmc_plot(fit.logst.rain.all, type = "trace")
mcmc_plot(fit.NB.rain.all, type = "trace")


```

```{r}
summary(fit.logst.rain.all, type = "trace")
summary(fit.NB.rain.all, type = "trace")


```













 