---
title: "Plots"
author: "Lisa Luna"
date: "9/22/2021"
output: html_document
---

This notebook creates figures 2, 3, S1, and S2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Load libraries


library(dplyr)
library(modelr)
library(ggdist)
library(tidybayes)
library(ggplot2)
library(cowplot)
library(brms)
library(RColorBrewer)
library(boot)
library(tibble)
library(patchwork)

theme_set(theme_tidybayes() + panel_border())

datadir <- '../01_Data/'
figdir <- '../03_Figures/'

```

```{r}
#read monthly precipitation summaries for comparison 

precip <- read.csv(paste0(datadir, 'footprint_precip.csv'))

```


# Figure 1

Made in python.  See Jupyter Notebook.


# Figure 2

Posterior parameter estimates for negative binomial models and logistic models - multi-inventory results alongside single-inventory results 

```{r}

#extract draws for each inventory from multi-inventory model

nbmeanML <- fit_all_NB %>%
    
          spread_draws(b_Intercept, r_month[month,], r_inv[inventory,]) %>%
          
          median_hdi(month_mean = exp(b_Intercept + r_month + r_inv), .width = c(0.95)) %>%
        
          mutate(Model = "Multi-inventory")



#extract draws from single-inventory Models

#nasa

nbmeannasa <- fit_nasa_NB %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_mean = exp(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "nasa", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


#slidoh

nbmeanslidoh <- fit_slidoh_NB %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_mean = exp(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "slidoh", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")

         
#slidod

nbmeanslidod <- fit_slidod_NB %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_mean = exp(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "slidod", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")

#seattle

nbmeanseattle <- fit_seattle_NB %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_mean = exp(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "seattle", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


#wlc

nbmeanwlc <- fit_wlc_NB %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_mean = exp(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "wlc", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


# combine into single dataframe 

nbmeandf <- bind_rows(nbmeanML, nbmeannasa, nbmeanslidoh, nbmeanslidod, nbmeanseattle, nbmeanwlc)

#clean up variable space
rm(nbmeanML, nbmeannasa, nbmeanslidoh, nbmeanslidod, nbmeanseattle, nbmeanwlc)

#do some variable name clean up for nicer plotting

nbmeandf[nbmeandf == "nasa"] <- "NASA GLC"
nbmeandf[nbmeandf == "slidoh"] <- "SLIDOh"
nbmeandf[nbmeandf == "slidod"] <- "SLIDOd"
nbmeandf[nbmeandf == "seattle"] <- "Seattle"
nbmeandf[nbmeandf == "wlc"] <- "WLC"

```

Negative Binomial Mean

```{r}

ordinv = c("NASA GLC", "SLIDOh", "SLIDOd", "Seattle", "WLC")

#plot posterior parameter estimates for the negative binomial mean from multi-inventory and single-inventory models for each inventory

nbmean <- nbmeandf %>%
        
        
        ggplot(aes(x = as.factor(month), y = month_mean, ymin = .lower, ymax = .upper, color = Model, shape = Model)) +
                  
                  geom_pointinterval(position = position_dodge(width = 0.6), 
                                     point_size = 1.8, 
                                     interval_size = 1.2) +
  
                  scale_color_manual(values = c("#005a32", "#999999")) +
                  
                  scale_y_log10() + 
  
                  coord_cartesian(ylim = c(10^-3, 10^3.5)) + 
                    
                  facet_wrap(~factor(inventory, levels = ordinv), nrow = 5) + 
  
                 scale_x_discrete(breaks=as.factor((1:12)),
                                        labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +
                 
                 ggtitle("A. Intensity") + 
  
                  ylab(expression(paste("Mean landslide intensity (landslides/100,000 ", km^{2}, "/month)"))) + 
      
                  theme_minimal() + 
  
                 theme(panel.grid.minor.y = element_blank(), 
                       panel.background = element_rect(fill = NA, color = 'black'), 
                       strip.text = element_blank(), 
                       axis.title.x = element_blank(), 
                       plot.title = element_text(face = "bold", size = 13), 
                       axis.title.y = element_text(size = 10), 
                       legend.title = element_blank(), 
                       legend.position = "none")
  
              


plot(nbmean)

```

Negative Binomial Variance

the variance of a negative binomially distributed random variable is mu + (mu^2/phi)

```{r}

#negative binomial variance estimates from multi-level model

nbvarML <-   fit_all_NB %>%
  
           spread_draws(b_Intercept, r_month[month,], r_inv[inventory,], #to calculate mu
                        b_shape_Intercept, r_month__shape[month,], r_inv__shape[inventory,]) %>% #to calculate phi

           median_hdi(month_var = exp(b_Intercept + r_month + r_inv) + 
                        (exp(b_Intercept + r_month + r_inv))^2/exp(b_shape_Intercept + r_month__shape + r_inv__shape),
                      .width = c(0.95)) %>% 
                     
              mutate(Model = "Multi-inventory")


#from separate models 

#nasa

nbvarnasa <- fit_nasa_NB %>%
  
              spread_draws(b_Intercept, r_month[month,], #to calculate mu
                        b_shape_Intercept, r_month__shape[month,]) %>% #to calculate phi

              median_hdi(month_var = exp(b_Intercept + r_month) + 
                        (exp(b_Intercept + r_month))^2/exp(b_shape_Intercept + r_month__shape),
                      .width = c(0.95))  %>%
  
              mutate(inventory = "nasa", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


#slidoh

nbvarslidoh <- fit_slidoh_NB %>%
  
              spread_draws(b_Intercept, r_month[month,], #to calculate mu
                        b_shape_Intercept, r_month__shape[month,]) %>% #to calculate phi

              median_hdi(month_var = exp(b_Intercept + r_month) + 
                        (exp(b_Intercept + r_month))^2/exp(b_shape_Intercept + r_month__shape),
                      .width = c(0.95))  %>%
  
              mutate(inventory = "slidoh", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")

         
#slidod

nbvarslidod <- fit_slidod_NB %>%
  
              spread_draws(b_Intercept, r_month[month,], #to calculate mu
                        b_shape_Intercept, r_month__shape[month,]) %>% #to calculate phi

              median_hdi(month_var = exp(b_Intercept + r_month) + 
                        (exp(b_Intercept + r_month))^2/exp(b_shape_Intercept + r_month__shape),
                      .width = c(0.95))  %>%
  
              mutate(inventory = "slidod", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")

#seattle

nbvarseattle <- fit_seattle_NB %>%
  
              spread_draws(b_Intercept, r_month[month,], #to calculate mu
                        b_shape_Intercept, r_month__shape[month,]) %>% #to calculate phi

              median_hdi(month_var = exp(b_Intercept + r_month) + 
                        (exp(b_Intercept + r_month))^2/exp(b_shape_Intercept + r_month__shape),
                      .width = c(0.95))  %>%
  
              mutate(inventory = "seattle", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


#wlc

nbvarwlc <- fit_wlc_NB %>%
  
              spread_draws(b_Intercept, r_month[month,], #to calculate mu
                        b_shape_Intercept, r_month__shape[month,]) %>% #to calculate phi

              median_hdi(month_var = exp(b_Intercept + r_month) + 
                        (exp(b_Intercept + r_month))^2/exp(b_shape_Intercept + r_month__shape),
                      .width = c(0.95))  %>%
  
              mutate(inventory = "wlc", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


# combine into single dataframe 

nbvardf <- bind_rows(nbvarML, nbvarnasa, nbvarslidoh, nbvarslidod, nbvarseattle, nbvarwlc)

rm(nbvarML, nbvarnasa, nbvarslidoh, nbvarslidod, nbvarseattle, nbvarwlc)

#do some variable name clean up for nicer plotting

nbvardf[nbvardf == "nasa"] <- "NASA GLC"
nbvardf[nbvardf == "slidoh"] <- "SLIDOh"
nbvardf[nbvardf == "slidod"] <- "SLIDOd"
nbvardf[nbvardf == "seattle"] <- "Seattle"
nbvardf[nbvardf == "wlc"] <- "WLC"


```

```{r}

#plot posterior parameter estimates for the negative binomial variance from multi-level and separate models for each inventory

nbvar <- nbvardf %>%
        
        
        ggplot(aes(x = as.factor(month), y = month_var, ymin = .lower, ymax = .upper, color = Model, shape = Model)) +
                  
                  geom_pointinterval(position = position_dodge(width = 0.6), 
                                     point_size = 1.8, 
                                     interval_size = 1.2) +
  
                  scale_color_manual(values = c("#238b45", "#999999")) + 
                  
                  scale_y_log10() + 
  
                  coord_cartesian(ylim = c(10^-3, 10^7)) + 
                    
                  facet_wrap(~factor(inventory, levels = ordinv), nrow = 5) + #, labeller = facetnames) + 
  
                  scale_x_discrete(breaks=as.factor((1:12)),
                                        labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +  
          
                  ggtitle("B. Variance") + 
  
                  ylab(expression(paste("Variance (landslides/100,000 ", km^{2}, "/month)"))) + 
      
                  theme_minimal() + 
  
                 theme(panel.grid.minor.y = element_blank(), 
                       panel.background = element_rect(fill = NA, color = 'black'), 
                       strip.text = element_blank(), 
                       axis.title.x = element_blank(), 
                       plot.title = element_text(face = "bold", size = 13), 
                       axis.title.y = element_text(size = 10), 
                       legend.position = "none")
  
                
  
                


plot(nbvar)


```

Logistic regression parameter estimates

```{r}

#extract draws for each inventory from multi-level model

logstpML <- fit_all_logst %>%
    
          spread_draws(b_Intercept, r_month[month,], r_inv[inventory,]) %>%
  
            median_hdi(month_p = inv.logit(b_Intercept + r_month + r_inv), .width = c(0.95)) %>%
        
          mutate(Model = "Multi-inventory")



#extract draws for each inventory from each single-inventory Model

#nasa

logstpnasa <- fit_nasa_logst %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_p = inv.logit(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "nasa", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


#slidoh

logstpslidoh <- fit_slidoh_logst %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_p = inv.logit(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "slidoh", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")

         
#slidod

logstpslidod <- fit_slidod_logst %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_p = inv.logit(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "slidod", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")

#seattle

logstpseattle <- fit_seattle_logst %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_p = inv.logit(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "seattle", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


#wlc

logstpwlc <- fit_wlc_logst %>%
  
              spread_draws(b_Intercept, r_month[month,]) %>%
  
              median_hdi(month_p = inv.logit(b_Intercept + r_month), .width = c(0.95)) %>%
  
              mutate(inventory = "wlc", .after = "month") %>% 
                     
              mutate(Model = "Single-inventory")


# combine into single dataframe 

logstpdf <- bind_rows(logstpML, logstpnasa, logstpslidoh, logstpslidod, logstpseattle, logstpwlc)

rm(logstpML, logstpnasa, logstpslidoh, logstpslidod, logstpseattle, logstpwlc)

#do some variable name clean up for nicer plotting

logstpdf[logstpdf == "nasa"] <- "NASA GLC"
logstpdf[logstpdf == "slidoh"] <- "SLIDOh"
logstpdf[logstpdf == "slidod"] <- "SLIDOd"
logstpdf[logstpdf == "seattle"] <- "Seattle"
logstpdf[logstpdf == "wlc"] <- "WLC"


```


```{r}
#plot posterior parameter estimates from the logistic regression from multi-inventory and single-inventory models for each inventory

logstp <- logstpdf %>%
        
        
        ggplot(aes(x = as.factor(month), y = month_p, ymin = .lower, ymax = .upper, color = Model, shape = Model)) +
                  
                  geom_pointinterval(position = position_dodge(width = 0.6), 
                                     point_size = 1.8, 
                                     interval_size = 1.2) +
  
                  scale_color_manual(values = c("#3f007d", "#999999")) + 
                  
                 coord_cartesian(ylim = c(0, 1)) + 
                    
                  facet_wrap(~factor(inventory, levels = ordinv), nrow = 5, strip.position = "right") + 
  
                  scale_x_discrete(breaks=as.factor((1:12)),
                                        labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) +   
  
                ggtitle("C. Probability") +
  
                  ylab("Landslide probability") + 
      
                  theme_minimal() + 
  
                 theme(panel.grid.minor.y = element_blank(), 
                       panel.background = element_rect(fill = NA, color = 'black'), 
                       strip.text = element_text(size = 12, face = 'bold'),
                       strip.background = element_rect(fill = "#E8E8E8", color = 'black'), 
                       axis.title.x = element_blank(), 
                       plot.title = element_text(size = 13, face = "bold"), 
                       axis.title.y = element_text(size = 10)) +
  
                theme(legend.position = "none") 

#plot(logstp)



```

```{r}

fig2 <- nbmean + nbvar + logstp 

ggsave(paste0(figdir, "fig2.png"), plot = fig2, width = 190, height = 170, units = c("mm"), dpi = 300)
  
```


```{r}
nbmeandf <- arrange(nbmeandf, inventory, Model, month)

nbvardf <- arrange(nbvardf, inventory, Model, month)

logstpdf <- arrange(logstpdf, inventory, Model, month)

```


```{r}
#summary results table for negative binomial mu parameters 

write.csv(nbmeandf, file = paste0(figdir, "nbmeandf.csv"))


#summary results table for negative binomial variance 
write.csv(nbvardf, file = paste0(figdir, "nbvardf.csv"))

#summary results table for logistic regression parameter

write.csv(logstpdf, file = paste0(figdir, "logstpdf.csv"))

```

# Figure 3 - Multi-inventory posterior predictives and precipitation  

Plot posterior predictive for multi-inventory negative binomial

```{r}

#posterior predictive from multi-inventory model for each inventory

invnames <- c("nasa", "slidoh", "slidod", "seattle", "wlc")

#set up new data that we want to predict for

newdata = data.frame("month" = integer(), "ln_area" = double(), "inv" = character())

for (i in invnames) {
  
  tempdf = subset(slidesallmi, inventory == i) %>%
          data_grid(month, ln_area, inv = i)
  
  newdata = bind_rows(newdata, tempdf)
}

#posterior predictive 

negbinom_pp_MI <- newdata %>%
                  add_predicted_draws(fit_all_NB) %>%
                  mutate(inventory = inv)

```


```{r}
#drop the 4000 landslides point from SLIDOh to truncate the axis for visualization: 

ord <-c("nasa", "slidoh", "slidod", "seattle", "wlc")

plMIppNB <-  negbinom_pp_MI %>%
            ggplot(aes(y = .prediction, x = factor(month))) +
              stat_interval(.width = c(.50, .80, .95, 0.99)) +
              geom_point(data = slidesallmi[slidesallmi$counts < 4000,], aes(y = counts, x = factor(month)), size = 1, alpha = 1/2) +
              scale_color_brewer(palette = "Greens", direction = -1) +
   
               scale_x_discrete(breaks=as.factor((1:12)),
                                labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) + 
              
              facet_wrap(~factor(inventory, levels = ord), nrow = 5, scale = "free_y") + 
                      
                     ylab("Monthly landslides") + 
              
                    theme_minimal() + 
              
                    theme(panel.grid.minor.y = element_blank(), 
                          panel.grid.major.x = element_blank(),
                          panel.background = element_rect(fill = NA, color = 'black'),
                          strip.text = element_blank(),
                          axis.title.x = element_blank(),
                           axis.title.y = element_text(size = 10),
                           legend.position = "none") 
                         
```


Posterior predictive for multi-inventory logistic regression 

```{r}

#posterior predictive 

logst_pp_MI <- newdata %>%
                  add_predicted_draws(fit_all_logst)

#summarize (% of months where 1 or more landslides are predicted)

logst_pp_MI_sum <- logst_pp_MI %>%
      
                    group_by(inv, month) %>% 
              
                    summarize(ppmean = mean(.prediction)) %>%
  
                    mutate(ppmean = ppmean*100) %>%
    
                    mutate(event = "Landslide") %>%
      
                    mutate(inventory = inv)

```

```{r}
#summarize landslide observations for comparison with predictions 

#(% of months where 1 or more landslides are predicted)

slidesallmi_event_sum <- slidesallmi %>%
      
                    group_by(inventory, month) %>% 
              
                    summarize(ppmean = mean(event)) %>%
  
                    mutate(ppmean = ppmean*100) %>%
    
                    mutate(event = "Landslide")
      

```


```{r}

# Plot summary of posterior predictive for logistic regression 

plMIpplogst <-  logst_pp_MI_sum %>%
      
        ggplot(aes(y=ppmean, x=as.factor(month))) + 
        
        geom_bar(stat = "identity", fill = "#6a51a3") + 
  
        geom_point(data = slidesallmi_event_sum, 
                   mapping = aes(x = as.factor(month), y = ppmean), 
                   size = 1.5, 
                   shape = 18, 
                   color = "black", 
                   fill = "black") + 
  
         scale_x_discrete(breaks=as.factor((1:12)),
                                labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) + 
  
        coord_cartesian(ylim = c(0,100)) + 
      
        facet_wrap(~factor(inventory, levels = ord), nrow = 5) + 
  
        ylab("% of months with one or more landslides") +
  
               theme_minimal() + 
  
        theme(panel.grid.minor.y = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.background = element_rect(fill = NA, color = 'black'),
              strip.text = element_blank(),
              axis.title.x = element_blank(),
              axis.title.y = element_text(size = 10),
               legend.position = "none") 
             

```

Plot mean monthly precipitation over the footprint areas covered by the landslide inventories

```{r}
precip[precip == "nasa"] <- "NASA GLC"
precip[precip == "slidoh"] <- "SLIDOh"
precip[precip == "slidod"] <- "SLIDOd"
precip[precip == "seattle"] <- "Seattle"
precip[precip == "wlc"] <- "WLC"


pl_precip <-  precip %>%
  
            ggplot(aes(x = as.factor(month), y = mean)) +

            geom_bar(stat = "identity", fill = '#08519c', width = 0.2) + 
            
  
         scale_x_discrete(breaks=as.factor((1:12)),
                                labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) + 
      
        facet_wrap(~factor(inventory, levels = ordinv), nrow = 5, strip.position = "right") + 
  
        ylab("Mean monthly precipitation (mm)") + 
  
        theme_minimal() + 
  
        theme(panel.grid.minor.y = element_blank(), 
              panel.grid.major.x = element_blank(),
              panel.background = element_rect(fill = NA, color = 'black'),
              strip.text = element_text(size = 12, face = 'bold'),
              strip.background = element_rect(fill = "#E8E8E8", color = 'black'), 
              axis.title.x = element_blank(),
               axis.title.y = element_text(size = 10),
               legend.position = "none")

```



```{r}

fig3 <- plMIppNB |  plMIpplogst | pl_precip

fig3 <- fig3 + plot_annotation(tag_levels = c("A"))

```



```{r}
ggsave(paste0(figdir, "fig3.png"), plot = fig3, width = 190, height = 170, units = c("mm"), dpi = 300)

```


Figure S1 - group level standard deviations 

```{r}

#negative binomial mu - month (sigma1m) and inventory (sigma2v)

#"sd_inv__Intercept"   "sd_month__Intercept"    

 a <- fit_all_NB %>%
    
          spread_draws(sd_inv__Intercept) %>%
  
          rename(value = sd_inv__Intercept) %>%
  
        mutate(group = "inventory") %>% 
  
        mutate(param = "sigma")

b <- fit_all_NB %>%
    
          spread_draws(sd_month__Intercept) %>%
  
         rename(value = sd_month__Intercept) %>%
  
        mutate(group = "month") %>% 
  
        mutate(param = "sigma")

sigma <- bind_rows(a, b)

rm(a,b)
       

#negative binomial phi -  month (psi1m) and inventory (psi2v)


#"sd_inv__shape_Intercept"     "sd_month__shape_Intercept"

c <- fit_all_NB %>%
    
          spread_draws(sd_inv__shape_Intercept) %>%
  
          rename(value = sd_inv__shape_Intercept) %>%
  
        mutate(group = "inventory") %>% 
  
        mutate(param = "psi")

d <- fit_all_NB %>%
    
          spread_draws(sd_month__shape_Intercept) %>%
  
         rename(value = sd_month__shape_Intercept) %>%
  
        mutate(group = "month") %>% 
  
        mutate(param = "psi")

psi <- bind_rows(c, d)

rm(c,d)

# logistic p - month (tau1m) and inventory (tau2v)

e <- fit_all_logst %>%
    
          spread_draws(sd_inv__Intercept) %>%
  
          rename(value = sd_inv__Intercept) %>%
  
        mutate(group = "inventory") %>% 
  
        mutate(param = "tau")


f <- fit_all_logst %>%
    
          spread_draws(sd_month__Intercept) %>%
  
          rename(value = sd_month__Intercept) %>%
  
        mutate(group = "month") %>% 
  
        mutate(param = "tau")


tau <- bind_rows(e, f)

rm(e,f)

group_level_sd <- bind_rows (sigma, psi, tau)


```

```{r}

figS1 <-  group_level_sd %>%
        
          ggplot(aes(x = value, y = group)) + 
               
          stat_halfeye(point_interval = median_hdi, .width = c(0.95))  + 
        
          facet_wrap(~factor(param, levels = c("sigma", "psi", "tau")), labeller = label_parsed) + 
        
           xlab("Parameter estimate") + 
        
           ylab(element_blank()) + 
        
                xlim(c(0,7)) 


ggsave(paste0(figdir, "figS1.png"), plot = figS1, width = 190, height = 90, units = c("mm"), dpi = 300)
  

```


Figure S2 - posterior predictives from single-inventory models

Plot posterior predictive for single-inventory negative binomial

```{r}

#posterior predictive from single-inventory model for each inventory

invnames <- c("nasa", "slidoh", "slidod", "seattle", "wlc")

fits_SINB <- list(fit_nasa_NB, fit_slidoh_NB, fit_slidod_NB, fit_seattle_NB, fit_wlc_NB)

#loop through single inventory model fits 

for (i in 1:5) {
  
  tempdf = subset(slidesallmi, inventory == invnames[i]) %>%
          data_grid(month, ln_area, inv = invnames[i])
  
  tempdf <- tempdf %>% add_predicted_draws(fits_SINB[[i]])
  
  if (i == 1) {
    
    negbinom_pp_SI <- tempdf
    
  } else {
    
    negbinom_pp_SI <- bind_rows(negbinom_pp_SI, tempdf)
    
  }
}

negbinom_pp_SI <- negbinom_pp_SI %>% rename(inventory = inv)

```


```{r}
#drop the 4000 landslides point from SLIDOh to truncate the axis for visualization: 

ord <-c("nasa", "slidoh", "slidod", "seattle", "wlc")

plSIppNB <-  negbinom_pp_SI %>%
            ggplot(aes(y = .prediction, x = factor(month))) +
              stat_interval(.width = c(.50, .80, .95, 0.99)) +
              geom_point(data = slidesallmi[slidesallmi$counts < 4000,], aes(y = counts, x = factor(month)), size = 1, alpha = 1/3) +
              scale_color_brewer(palette = "Greens", direction = -1) +
   
               scale_x_discrete(breaks=as.factor((1:12)),
                                labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) + 
              
              facet_wrap(~factor(inventory, levels = ord), nrow = 5, scale = "free_y") + 
                      
                     ylab("Monthly landslides") + 
              
                    theme_minimal() + 
              
                    theme(panel.grid.minor.y = element_blank(), 
                          panel.grid.major.x = element_blank(),
                          panel.background = element_rect(fill = NA, color = 'black'),
                          strip.text = element_blank(),
                          axis.title.x = element_blank(),
                          axis.title.y = element_text(size = 10),
                          legend.position = "none") 
                         
```


Plot posterior predictive for single-inventory logistic 

```{r}


#posterior predictive from single-inventory model for each inventory

invnames <- c("nasa", "slidoh", "slidod", "seattle", "wlc")

fits_SIlogst <- list(fit_nasa_logst, fit_slidoh_logst, fit_slidod_logst, fit_seattle_logst, fit_wlc_logst)

#loop through single inventory model fits 

for (i in 1:5) {
  
  tempdf = subset(slidesallmi, inventory == invnames[i]) %>%
          data_grid(month, ln_area, inv = invnames[i])
  
  tempdf <- tempdf %>% add_predicted_draws(fits_SIlogst[[i]])
  
  if (i == 1) {
    
    logst_pp_SI <- tempdf
    
  } else {
    
    logst_pp_SI <- bind_rows(logst_pp_SI, tempdf)
    
  }
}

#summarize (% of months where 1 or more landslides are predicted)

logst_pp_SI_sum <- logst_pp_SI %>%
      
                    group_by(inv, month) %>% 
              
                    summarize(ppmean = mean(.prediction)) %>%
  
                    mutate(ppmean = ppmean*100) %>%
    
                    mutate(event = "Landslide") %>%
      
                    rename(inventory = inv)


```


```{r}

# Plot summary of posterior predictive for logistic regression 

plSIpplogst <-  logst_pp_SI_sum %>%
      
        ggplot(aes(y=ppmean, x=as.factor(month))) + 
        
        geom_bar(stat = "identity", fill = "#6a51a3") + 
      
        geom_point(data = slidesallmi_event_sum, 
                         mapping = aes(x = as.factor(month), y = ppmean), 
                         size = 1.5, 
                         shape = 18, 
                         color = "black", 
                         fill = "black") + 
  
         scale_x_discrete(breaks=as.factor((1:12)),
                                labels=c("J", "F", "M", "A", "M", "J", "J", "A", "S", "O", "N", "D")) + 
  
        coord_cartesian(ylim = c(0,100)) + 
      
        facet_wrap(~factor(inventory, levels = ord), nrow = 5) + 
  
        ylab("% of months with one or more landslides") +
  
               theme_minimal() + 
  
        theme(panel.grid.minor.y = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.background = element_rect(fill = NA, color = 'black'),
              strip.text = element_blank(),
              axis.title.x = element_blank(),
              axis.title.y = element_text(size = 10),
              legend.position = "none") 
             

```


```{r}

figS2 <- plSIppNB |  plSIpplogst | pl_precip

figS2 <- figS2 + plot_annotation(tag_levels = c("A"))

```



```{r}
ggsave(paste0(figdir, "figS2.png"), plot = figS2, width = 190, height = 170, units = c("mm"), dpi = 300)

```

